---
title: "COVID-19-analysis"
author: "Hubert Krzyżanowski"
date: "12 11 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(tidyr)
# Read .xlsx to dataframe
library(readxl)
# Change seconds to period
library(lubridate)
# Data visualization
library(ggplot2)
# Table design
library(kableExtra)
# Interactive plot
library(plotly)
# Plot animation - gifs
library(gganimate)
# Allocate function
library(purrr)
# User friendly data summary tbl_summary
library(gtsummary)
# Dataset split for classification problem
library(caTools)
# CARET - machine learning algorithms
library(caret)
```

```{r data_reading}
org_df <- read_excel("wuhan_blood_sample_data_Jan_Feb_2020.xlsx")

```
### Data cleaning
The following steps were undertaken to clean the original dataset:
<ul>
  <li> replace the _gender_ values, from numeric (1, 2) to factor (male, female),
  <li> replace the _outcome_ values, from numeric (0,1) to factor (survived, died),
  <li> unification of column name from _'Admission time'_ to _admission_time_,
  <li> unification of column name from _'Discharge time'_ to _discharge_time_,
  <li> replace _NA_ values in _PATIENT_ID_ with suitable id (done in next chapter).
</ul>

```{r dataset_cleaning}
df <- org_df %>% 
        mutate(gender = as.factor(ifelse(gender==1, "male", "female"))) %>%
        mutate(outcome = as.factor(ifelse(outcome == 0, "survived", "died"))) %>%
        rename(admission_time = 'Admission time',
               discharge_time = 'Discharge time')

```

The dataset was split into two dataframes in order to make data analysis easier.

#### Patients
One additional column was created to store the hospitalization time of all patients - used in further analysis to check the relation between length stay and outcome. 

```{r create_patients_df, cache=TRUE}
patients <- df %>% 
              select(PATIENT_ID, age, gender, admission_time, discharge_time, outcome) %>% 
              drop_na(PATIENT_ID) %>%
              mutate("hospitalization_length" = round((difftime(discharge_time, admission_time, units = "days") ), digits = 2)) %>%
              relocate(hospitalization_length, .after = discharge_time)

head(patients) %>%
  kbl() %>%
  kable_paper("hover")

```
#### Blood tests

```{r blood_tests_df, cache= TRUE}
blood_tests_df <- df %>%
  select(-c(admission_time, discharge_time)) %>%
  fill(PATIENT_ID)

blood_tests_df %>% select (-PATIENT_ID) %>%
                   tbl_summary(by = outcome, missing = "no", statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
                   modify_header(label = "**Variable**")
```

```{r patients_plot_gender}
ggplot(patients, aes(x = gender, fill = gender)) +
  geom_bar() + 
  labs(title= "Numer of patients per gender hospitilized in Tongji Hospital (Wuhan) ", 
       subtitle = "between 10 January and 18 February 2020",
       y = "Number of patients", 
       x = "Gender")
```

## Visualization

### Patients grouped by gender

```{r patients_plot_hist, fig.width= 10}
patients_hist <- ggplot(patients, aes(x = age, fill = gender)) +
  geom_histogram(stat = "count",
                 binwidth = 1.2)+
  labs(y = "Number of patients", 
       x = "Age") +
  scale_x_continuous(breaks=seq(20, 100, 5))
 
ggplotly(patients_hist)      

```


### Outcome grouped by gender, age

```{r out_per_gender_plot, fig.width= 10}

patients_outcome <- ggplot(patients, aes(x = age, fill = outcome)) + 
                    geom_histogram(binwidth = 1.2) +
                    facet_grid(~ gender) +
                    scale_y_continuous(breaks=seq(0, 20, 2)) +
                    scale_x_continuous(breaks=seq(20, 100, 5)) + ylab("Number of cases")

ggplotly(patients_outcome)
```

### Outcome due to hospitalization length grouped by gender
```{r hospitalization_plot, fig.width= 9}

ggplot(patients, aes(x = hospitalization_length, fill = outcome)) + 
      geom_histogram(binwidth = 1.2) +
      facet_grid(outcome ~ gender) +
      scale_y_continuous(breaks=seq(0, 20, 2)) +
      scale_x_continuous(breaks=seq(0, 40, 5)) +
      labs(y = "Number of patients", 
           x = "Hospitalization length [days]")

```

### Outcome animation over time
```{r animation_plot, fig.width= 9}

```

### Relation between outcome dead and biomarker

```{r biomarker}


```

### Classification problem

```{r }

```

```{r dataset_split}
# set.seed(100)
# 
# training_index <- createDataPartition(blood_tests_df$outcome, p = 0.7, list = FALSE)
# train <- blood_tests_df[training_index,]
# test <- blood_tests_df[-training_index,]
# # 
# # # SVM Model
# # # Build training model
# # model <- train(outcome ~ ., data = blood_tests_df,
# #                method = "svmPoly",
# #                na.action = na.omit,
# #                preProcess = c("scale","center"),
# #                trControl = trainControl(method = "none"),
# #                tuneGrid = data.frame(degree = 1, scale = 1, C = 1))
# # 
# # # Build CV Model
# model_cv <- train(outcome ~ ., data = blood_tests_df,
#                method = "svmPoly",
#                na.action = na.omit,
#                preProcess = c("scale","center"),
#                trControl = trainControl(method = "cv", number = 10),
#                tuneGrid = data.frame(degree = 1, scale = 1, C = 1))
# # 
# model_cv <- predict(model_cv, train)
# 
# model_cv_confusion <- confusionMatrix(model_cv, test$outcome)
```