---
title: "COVID-19-analysis"
author: "Hubert Krzyżanowski"
date: "12.11.2020"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r libraries}
library(knitr)
library(dplyr)
library(tidyr)
# Read .xlsx to dataframe
library(readxl)
# Change seconds to period
library(lubridate)
# Data visualization
library(ggplot2)
# Table design
library(kableExtra)
# Interactive plot
library(plotly)
# Plot animation - gifs
library(gganimate)
# Allocate function
library(purrr)
# User friendly data summary tbl_summary
library(gtsummary)
# Dataset split for classification problem
library(caTools)
# CARET - machine learning algorithms
library(caret)
# Replacing NA values in column
library(imputeTS)
# Correlate function
library(corrr)
# ROC Curve
library(pROC)
# Correlaion plots
library(corrplot)
# Grid arrange
library(gridExtra)
```

```{r data_reading}
org_df <- read_excel("wuhan_blood_sample_data_Jan_Feb_2020.xlsx")
```

# __Dataset cleaning__
The following steps were undertaken to clean the original dataset:
<ul>
  <li> replace the _gender_ values, from numeric (1, 2) to factor (male, female),
  <li> replace the _outcome_ values, from numeric (0,1) to factor (survived, Died),
  <li> unification of column name from _'Admission time'_ to _admission_time_,
  <li> unification of column name from _'Discharge time'_ to _discharge_time_,
  <li> replace _NA_ values in _PATIENT_ID_ with suitable id (done in next chapter),
  <li> removing patients with no biomarker values at all,
  <li> renaming biomarker columns.
</ul>

```{r dataset_cleaning}
df <- org_df %>% 
        mutate(gender = as.factor(ifelse(gender==1, "male", "female"))) %>%
        mutate(outcome = as.factor(ifelse(outcome == 0, "Survived", "Died"))) %>%
        filter(!is.na(org_df$RE_DATE)) %>% 
        rename(admission_time = 'Admission time',
               discharge_time = 'Discharge time',
               hs_CRP = 'High sensitivity C-reactive protein')

names(df)[34] <- "Tumor necrosis factor alpha"
names(df)[37] <- "Interleukin 1 beta"
names(df)[68] <- "Gamma glutamyl transpeptidase"

```

 
# __Dataset summary__
The dataset consists of `r ncol(df)` variables and has `r nrow(df)` observations (blood tests). See the summary below:

```{r dataset_summary}
summary_df <- df %>% select(outcome, gender)
tbl_summary(
  summary_df,
  by = outcome,
  label = gender ~ "Gender") %>% 
  modify_header(label ~ "**Variable**") %>% 
  add_overall() %>%
  as_kable() %>%   kable_paper("hover")

```
 

The blood tests were taken from `r nrow(df %>% select(PATIENT_ID) %>% drop_na)` different patients.

```{r patients_summary}
df %>% select(PATIENT_ID, gender, outcome) %>% 
  drop_na(PATIENT_ID) %>% 
  select(-PATIENT_ID) %>%  
  tbl_summary(label =  gender ~ "Gender", by = outcome) %>%
  add_overall() %>% 
  modify_header(label ~ "**Variable**") %>%
  as_kable() %>%   kable_paper("hover")

```

 

**From the cleaned dataset, two dataset are created _Patients_ and _Blood tests_ containg specific values, in order to make data analysis easier.**



### Patients
One additional column was created to store the hospitalization time of all patients - used in further analysis to check the relation between length stay and outcome. Go to section [Patients Visualization](#visualization) to see basic visualizations about patients.

```{r create_patients_df}
patients <- df %>% 
              select(PATIENT_ID, age, gender, admission_time, discharge_time, outcome) %>% 
              drop_na(PATIENT_ID) %>%
              mutate("hospitalization_length" = round((difftime(discharge_time, admission_time, units = "days") ), digits = 2)) %>%
              relocate(hospitalization_length, .after = discharge_time)

head(patients) %>%
  kbl() %>%
  kable_paper("hover")

```

### Blood tests


```{r blood_tests_df}

blood_tests_df <- df %>%
  select(-c(admission_time, discharge_time)) %>%
  fill(PATIENT_ID)

markers_df <- blood_tests_df %>% select (-c(PATIENT_ID, age, RE_DATE, gender))

  tbl_summary(
    markers_df,
    by = outcome,
    missing = "no") %>% 
    modify_header(label = "**Marker**") %>%
    add_n() %>%
    bold_labels() %>%
    as_kable() %>%  
    kable_paper("hover") %>% 
    scroll_box(width = "100%", height = "200px")
```
The blood tests are prepared for further analysis. For each patient there were many blood samples, containing many missing values. All the samples have been combined into one sample containing the last value (closest to discharge). 

```{r last_sample_prepare}
last_sample_df <- blood_tests_df %>% 
  select(-RE_DATE) %>%
  group_by(PATIENT_ID) %>% 
  summarise(across(everything(), function(x) last(na.omit(x)))) %>%
  select(-PATIENT_ID)
```

The combined blood samples dataset was also preprocessed for classification. Go to section [Classification - dataset cleaning](#dataset-cleaning-1) to see how it was cleaned. Columns and patients with too many missing values were deleted from the dataset.

```{r class_df}
# %>% na_mean(option = "median")
class_df <- last_sample_df 
```
# __Visualization__

## Patients  gender 
```{r patients_plot_gender, fig.width= 9}
ggplot(patients, aes(x = gender, fill = gender)) +
  geom_bar() + 
  labs(y = "Number of patients", 
       x = "Gender") +
  theme(legend.position = "none")
```

## Patients grouped by age gender

```{r patients_plot_hist, fig.width= 9}
patients_hist <- ggplot(patients, aes(x = age, fill = gender)) +
  geom_histogram(stat = "count",
                 binwidth = 1.2)+
  labs(y = "Number of patients", 
       x = "Age") +
  scale_x_continuous(breaks=seq(20, 100, 5))
 
ggplotly(patients_hist)      

```

## Outcome grouped by gender, age

```{r out_per_gender_plot, fig.width= 9}
layout_ggplotly <- function(gg, x = -0.05, y = -0.05){
  # The 1 and 2 goes into the list that contains the options for the x and y axis labels respectively
  gg[['y']][['layout']][['annotations']][[1]][['y']] <- x
  gg[['y']][['layout']][['annotations']][[2]][['x']] <- y
  gg
}

patients_outcome <- ggplot(patients, aes(x = age, fill = outcome)) + 
                    geom_histogram(binwidth = 1.2) +
                    facet_grid(~ gender) +
                    scale_y_continuous(breaks=seq(0, 20, 2)) +
                    scale_x_continuous(breaks=seq(20, 100, 5)) +
                    labs(y = "Number of patients", x = "Age")

ggplotly(patients_outcome)
```

## Outcome due to hospitalization length grouped by gender
```{r hospitalization_plot, fig.width= 9}

hospitalization_length_plot <- ggplot(patients, aes(x = hospitalization_length, fill = outcome)) + 
      geom_histogram(binwidth = 1.2) +
      facet_grid(outcome ~ gender) +
      scale_y_continuous(breaks=seq(0, 20, 2)) +
      scale_x_continuous(breaks=seq(0, 40, 5)) +
      labs(y = "Number of patients", 
           x = "Hospitalization length [days]")

ggplotly(hospitalization_length_plot)

```

## Died in specific days

```{r outcome_per_day, fig.width= 9}

outcome_per_day <- patients %>% 
                    mutate(discharge_time = as.Date(discharge_time)) %>%
                    filter(outcome == "Died")
                  
outcome_per_day_plot <- ggplot(outcome_per_day, aes(x = discharge_time, fill = outcome)) + 
  geom_histogram(binwidth = 1.2) + 
  facet_grid(~ gender) +  
  labs(x = "Discharge date", y = "Number of deaths") + 
  theme(legend.position = "none")

ggplotly(outcome_per_day_plot)

```

## Dead cases during the day

```{r outcome_during_day, fig.width= 9}
outcome_during_day_plot <- patients %>%  
  mutate(time_h_m = hms(format(patients$discharge_time, format = "%H:%M:%S"))) %>% 
  mutate(time_h_m = (hour(time_h_m) + minute(time_h_m)/60)) %>%
  filter(outcome == "Died") %>%
  ggplot(aes(x = time_h_m, fill = "blue")) + 
  geom_histogram(binwidth = 1.2) + 
  scale_x_continuous(breaks = seq(0, 24, by = 1)) + 
  labs(x = "Number of dead cases", y = "Time of the day")+
  theme(legend.position = "none")

ggplotly(outcome_during_day_plot)
```

# __Variables correlation__

Preparing the dataset for correlation (changing factor variables to numeric).

```{r corr_prepare}
cor_df <- last_sample_df %>%
            mutate(outcome = ifelse(last_sample_df$outcome == "Died", 1, 0)) %>%
            mutate(gender = ifelse(last_sample_df$gender == "male", 1, 0)) %>%
            rename(male = gender)

correlationMatrix <-  correlate(cor_df[sapply(cor_df, is.numeric)], use='pairwise.complete.obs')
```

## Age correlation
From the previous analysis, it is known that elderly people are more susceptible to die due to Covid-19. Below short summary, what biomarkers are highly correlated with age.

```{r age_correlation}

age_correlation <- correlationMatrix %>% 
  focus(age) %>% 
  mutate(age = abs(age)) %>%
  arrange(desc(age)) %>% 
  filter(rowname != "outcome") %>% head 

age_correlation %>% kbl() %>% kable_paper("hover")

```
The most correlated is _eGFR_ which is used to measure the the effectiveness of the work of the kidneys. Its hard to present a norm value, because this marker depends on many factors like gender, age, body mass, but some sources show that value above 90 is proper. Too low ,and too high value of GFR in some cases indicate kidney diseases which affect the blood filtration. 

Below chart presents the GFR value between patients in different age, grouped by outcome. It's analysis shows, that many elderly patients that died, had some abnormalities in the work of the kidneys.

```{r GFR_corr_plots, fig.width= 9}
ggplot(last_sample_df, aes(x = age, y = `eGFR`, color = outcome)) + 
  geom_point() + 
  theme(legend.position = c(0.9,0.9)) + 
  ylim(0 , 150)

```

The next two high correlated biomarkers are related to immune system. The values of _lymphocyte_ and _neutrophils_ show how strong the organism is and how well it fights with the disease.

Lymphocytes are cells responsible for protecting our body (by creating anitbodies) from viruses, bacteria and other disease causing factors. The norm value for an adult is between 15 - 40%. Lower lymphocytes levels means, that the body cannot fight the disease. The left chart below confirms, that elderly people have weaker immune system and it's hard for their organism to fight the disease. 

```{r lym_corr_plots, fig.width= 9}
plot1 <- ggplot(last_sample_df, aes(x = age, y = `(%)lymphocyte`, color = outcome)) + geom_point() + theme(legend.position = "none")
plot2 <- ggplot(last_sample_df, aes(x = `lymphocyte count`, y = `(%)lymphocyte`, color = outcome)) + 
  geom_point() + 
  theme(legend.position = c(0.8, 0.2)) +
  xlim(0,3.75)
grid.arrange(plot1, plot2, ncol=2)
```

Neutrophils are essential part of immune system - this cells search for pathogens in organisms and destroy them. High value of _neutrophils(%)_ results in many neutrophil cells in blood (right plot below), which means that a medical condition occurs in patients body and that the immune system fights it.

This correlation explains that elderly people are more vulnerable, and their immune systems need to produce more neutrophils to fight the pathogens than younger patients.
The left plot shows that some of the tested patients had some medical condition, due to increased amount of neutrophils. Adding the information about the outcome, confirms that elderly patients are more likely to die because of Covid-19.
```{r neutrophil_corr_plots, fig.width= 9}
plot1 <- ggplot(last_sample_df, aes(x = age, y = `neutrophils(%)`, color = outcome)) + geom_point() + theme(legend.position = "none")
plot2 <- ggplot(last_sample_df, aes(x = `neutrophils count`, y = `neutrophils(%)`, color = outcome)) + geom_point() + theme(legend.position = c(0.8, 0.2))
grid.arrange(plot1, plot2, ncol=2)
```

## Outcome correlation {.tabset .tabset-pills}

The following section is devoted to check the correlation between biomarkers and the outcome. 

The correlation matrix for the highest correlated variables and the numeric correlation values are shown below.

```{r outcome_corr_plot, fig.width= 9}
'%ni%' <- Negate('%in%')

outcome_cor <- correlationMatrix %>% 
  focus(outcome) %>% 
  mutate(outcome = abs(outcome)) %>%
  arrange(desc(outcome)) %>% 
  filter(`rowname` %ni% c('neutrophils(%)', 'neutrophils count')) %>% 
  mutate(outcome = round(outcome,2)) %>%
  filter(outcome > 0.5)
  

outcome_corr_df <- cor_df %>% select(c(outcome_cor$rowname, outcome)) 

outcome_cor_matrix <- cor(outcome_corr_df[sapply(outcome_corr_df, is.numeric)], use='pairwise.complete.obs')

corrplot(outcome_cor_matrix)
```

The previous sections contains the analysis about _lymphocytes_ and how important they are when fighting the disease, that's why they won't be considered in this section. 

```{r outcome_corr}
outcome_cor %>% kbl() %>% kable_paper("hover") 
```

Below in each tab are presented the values of each biomarkers (correlation > 0.65) for all the patients grouped by age and outcome. 
Analysis of theses data shows, that all the biomarkers are also somehow correlated with the age, because the biomarkers values for eldery are very often (in this 5 biomarkers) outstanding from the values for people less than 50 years. This statement is confirmed by the boxplots below every chart, preseting the distribution of the biomarkers grouped by age group (adult - less than 64 years, eldery - more than 64 years), gender and outcome.


```{r plotly_function}
layout_ggplotly <- function(gg, x = -0.02, y = -0.05){
  # The 1 and 2 goes into the list that contains the options for the x and y axis labels respectively
  gg[['x']][['layout']][['annotations']][[1]][['y']] <- x
  gg[['x']][['layout']][['annotations']][[2]][['x']] <- y
  gg
}

```


### Albumin
```{r albumin_crp_plot}
ggplot(last_sample_df, aes(x = age, y = `albumin`, color = outcome)) + geom_point()

albumin_plot <- last_sample_df %>% 
  mutate(age_group = as.factor(ifelse(last_sample_df$age < 64, 'adult', 'elderly'))) %>% 
  ggplot(aes(x= age_group, y = `albumin`, fill = gender)) +
  geom_boxplot(na.rm=TRUE) +  facet_grid(~outcome) + 
  labs(x = "Age group", y = "Albumin")

ggplotly(albumin_plot) %>% layout(boxmode = "group") %>% layout_ggplotly
```

### Prothrombin activity

```{r pt_crp_plot}
ggplot(last_sample_df, aes(x = age, y = `Prothrombin activity`, color = outcome)) + geom_point()

pt_plot <- last_sample_df %>% 
  mutate(age_group = as.factor(ifelse(last_sample_df$age < 64, 'adult', 'elderly'))) %>% 
  ggplot(aes(x= age_group, y = `Prothrombin activity`, fill = gender)) +
  geom_boxplot(na.rm=TRUE) +  facet_grid(~outcome) + 
  labs(x = "Age group", y = "Prothrombin activity")

ggplotly(pt_plot) %>% layout(boxmode = "group") %>% layout_ggplotly
```


### Hs-CRP

A norm value for hs-CRP is about 50. All values above that level indicate some kind of inflammation in the body. Many values on the first plot are much more above the norm level showing very strong inflammation which eventually (probably) contributed to the death.

```{r hs_crp_plot}

ggplot(last_sample_df, aes(x = age, y = hs_CRP, color = outcome)) + geom_point()

crp_plot <- last_sample_df %>%
  mutate(age_group = as.factor(ifelse(last_sample_df$age < 64, 'adult', 'elderly'))) %>%
  ggplot(aes(x= age_group, y = hs_CRP, fill = gender)) +
  geom_boxplot(na.rm=TRUE) +  facet_grid(~outcome) +
  labs(x = "Age group", y = "High sensitivity C-reactive protein")

ggplotly(crp_plot) %>% layout(boxmode = "group") %>% layout_ggplotly
```

### D-D dimer 

D-dimers are cells responsible for decomposition of a clot. Their high value mean that there was a blood clot in the organism. Sometimes it can be linked with myocardial infarction, pulmonary embolism which combined with Covid-19 symptoms can lead to death. 

```{r d_dimer_plot}

ggplot(last_sample_df, aes(x = age, y = `D-D dimer`, color = outcome)) + geom_point()

dimer_plot <- last_sample_df %>% 
  mutate(age_group = as.factor(ifelse(last_sample_df$age < 64, 'adult', 'elderly'))) %>% 
  ggplot(aes(x= age_group, y = `D-D dimer`, fill = gender)) +
  geom_boxplot(na.rm=TRUE) +  facet_grid(~outcome) + 
  labs(x = "Age group", y = "D-D dimer")

ggplotly(dimer_plot) %>% layout(boxmode = "group") %>% layout_ggplotly
```

### Lactate dehydrogenase 
LDH are cells, which are released into the blood, where other cells are damaged or destroyed. Higher LDH levels in blood may be a sign of tissue damage or disease and they can be related with pneumonia, which combined with Covid-19 symptoms can lead also to death.
```{r ldh_plot}

ggplot(last_sample_df, aes(x = age, y = `Lactate dehydrogenase`, color = outcome)) + geom_point()

ldh_plot <- last_sample_df %>% 
  mutate(age_group = as.factor(ifelse(last_sample_df$age < 64, 'adult', 'elderly'))) %>% 
  ggplot(aes(x= age_group, y = `Lactate dehydrogenase`, fill = gender)) +
  geom_boxplot(na.rm=TRUE) +  facet_grid(~outcome) + 
  labs(x = "Age group", y = "Lactate dehydrogenase")

ggplotly(ldh_plot) %>% layout(boxmode = "group") %>% layout_ggplotly
```

## {-}

# __Animation__

Animated aggregate number of deaths in next days is presented below. A shoot up can be noticed between 02.02.2020 - 22.02.2020. After that the deaths levelled off, and another peak occcured on 04.04.2020.

```{r animation_plot}

patients_agg <- patients %>% select(c(discharge_time, outcome)) %>%
  mutate(discharge_time = as.Date(patients$discharge_time, "%m/%d/%Y" )) %>%
  filter(outcome == 'Died') %>%
  group_by(discharge_time) %>%
  summarise(deaths_count = n(), .groups="drop") %>%
  arrange(discharge_time) %>%
  mutate(deaths_count_agg = cumsum(deaths_count))

ggplot(patients_agg, aes(x = discharge_time, y = deaths_count_agg)) + 
  geom_line(size  = 1.1, color = 'red') + 
  transition_reveal(discharge_time) + 
  labs(x = "Discharde time", y = "Deaths count aggregate") + 
  scale_x_continuous(breaks = seq(min(patients_agg$discharge_time), max(patients_agg$discharge_time), 10))


```

# __Classification model__
In this chapter classification model is trained to predict the outcome (death/survival) of COVID-19 sick patients based on basic patients observations and blood test samples. One blood test for each patient is considered as an observation for the machine learning algorithm. As it was explained extra data pre processing was needed to prepare the dataset. For each patient, all the blood test are reduced to one row, containing the closest value to the discharge time.

Redundant columns like patient id, blood test time and admission and discharge time were removed from the dataset. 

## Dataset cleaning

For machine learning process there should be no missing values in the dataset. Summary below shows, that there are columns with many missing values.

```{r ml_dataset_prepare}

class_df %>% select(-c(age,gender, outcome)) %>% summary %>% kbl %>% kable_paper("hover") %>% scroll_box(width = "100%", height = "300px")

```

Below cleaning is done, to check if the dataset contains patients with basic info like age and gender, but with no many missing biomarker values - these patients are removed from the dataset.


```{r ml_dataset_prepare_}
#Deleting rows with no many missing values
rows_to_delete <- c()

for(i in 1:nrow(class_df)) {
  row_na_sum <- sum(is.na(class_df[i,]))
  if (row_na_sum >= 35) {
    rows_to_delete <- c(rows_to_delete, i)
  }
}
patients_to_delete <- length(rows_to_delete)

class_df <- class_df[-rows_to_delete, ]

class_df %>% select(-c(age,gender, outcome)) %>% summary %>% kbl %>% kable_paper("hover") %>% scroll_box(width = "100%", height = "300px")
```

`r patients_to_delete` patients are removed from the dataset, because they contain more than 40 missing values. 

Many columns have more than 70 missing values - they won't be used for the classification.

```{r ml_dataset_prepare_2}

class_df <- class_df %>% select(-c(`Interleukin 2 receptor`, `Interleukin 10`, `antithrombin`, `Interleukin 8`, `Quantification of Treponema pallidum antibodies`, `HBsAg`, `Tumor necrosis factor alpha`, `Interleukin 1 beta`, `PH value`, `ferritin`, `Amino-terminal brain natriuretic peptide precursor(NT-proBNP)`, `Interleukin 6` , `Fibrin degradation products`, `2019-nCoV nucleic acid detection`, `HIV antibody quantification`, `Hypersensitive cardiac troponinI`, `HCV antibody quantification`, `ESR`))

class_df %>% select(-c(age,gender, outcome)) %>% summary %>% kbl %>% kable_paper("hover") %>% scroll_box(width = "100%", height = "300px")

```

The remaining values in the dataset are replaced with median for the whole column - due to very skewed distribution. The summary of clean dataset, with no missing values is presented below.

```{r ml_dataset_prepare_3}

class_df <- class_df %>% na_mean(option = "median")

class_df %>% select(-c(age,gender, outcome)) %>% summary %>% kbl %>% kable_paper("hover") %>% scroll_box(width = "100%", height = "300px")
```

## Dataset shuffle and split

The preprocessed data is split into two datasets _training_ and _testing_.

The patients are grouped by outcome, first in the dataset are patients who survived and than those who died. Below dataset shuffle and check is done to be sure that the training and testing sets have similar output class distribution.

To ensure  the repeatability of experiments, seed is set to 23.

```{r class_data_split}
set.seed(23)
rows <- sample(nrow(class_df))
class_df <- class_df[rows,]


set.seed(23)
inTraining <- createDataPartition(y = class_df$outcome, p=.70, list=FALSE)
training <- class_df[inTraining,]
testing <- class_df[-inTraining,]
```


__Training set summary__
```{r ml_test_datasets}
training %>% select(gender, outcome) %>% tbl_summary(by = outcome) %>%  as_kable()  %>% kable_paper("hover")
```
__Testing set summary__
```{r ml_test_datasets_2}
testing %>% select(gender, outcome) %>% tbl_summary(by = outcome) %>%  as_kable()  %>% kable_paper("hover")
```

## Train control

For the learning process Repeated 2 fold Cross-Validation was used - the training process will be repeated 5 times.

```{r ml_train_ctrl}
set.seed(23)
ctrl <- trainControl(
    method = "repeatedcv",
    number = 2,
    repeats = 5,
    classProbs = TRUE)

```

Te measure the performance of the model three measures are considered: accuracy, ROC curve, and AUC.

## Random Forest

### Model training
The Random Forest model is trained with default parameters, but with a number of trees in the forest set to 10 and metric used for tuning the model as ROC. 

```{r ml_blood_test, warning=FALSE}

rfGrid <- expand.grid(mtry = 10:20)

set.seed(23)
rf_fit <- train(outcome ~ ., 
                data = training, 
                method = "rf",
                preProc = c("center", "scale"),
                trControl = ctrl,
                tuneGrid = rfGrid,
                ntree = 15)

rf_fit
```

### Prediction
The model works very well with 97% accuracy. There are 3 badly classified patients, but this type of error is less harmful (FN).

```{r ml_rf_predict}
rf_classes <- predict(rf_fit, newdata = testing)
rf_classes_prob <- predict(rf_fit, newdata = testing, type = "prob")
caret::confusionMatrix(data = rf_classes, testing$outcome)
```


### ROC and AUC
Presented ROC curve is very convex which means that the model works very well. Also the AUC value is very high, close to 1. 
```{r ml_rf_roc, warning= FALSE}
rf_ROC <- roc(response = testing$outcome, 
              predictor = rf_classes_prob[, "Died"],
              levels = rev(levels(testing$outcome)),
              plot = TRUE,
              auc = TRUE,
              print.auc = TRUE)
rf_ROC
```

### Feauture importance

Presented below classification variables and their importance show, that there are just few very important, decisive variable which model uses. 
Three variables: LDH, lymphocyte, and hs-CRP, are marked as the most important variables to predict the mortality of Covid-19 patients - the same as in the article [An interpretable mortality prediction model for COVID-19 patients](https://www.nature.com/articles/s42256-020-0180-7).
The analysis of this biomarkers was conducted in previews section - [Variables correlation](#variables-correlation)

```{r feature_importance}

importance <- varImp(rf_fit)

importance

```

## Further development

As further work development, the most important variables (importance > 5) could be used to train the model to get 100% classification accuracy. 
Extra model evaluation could be done with learning curves to detect overfitting.
